\pdftitle

### 2019-01-24 maxNt_v3.tex
that removed
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-11-04T09:20:08+0100
%% Last-Updated: 2018-07-19T11:05:41+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Report-no: ***
\documentclass{article}

@@

 \DeclareMathSymbol{\iota}{\mathalpha}{egreeki}{"13}

@@

\newcommand*{\p}{\mathrm{P}}%probability

@@

\newcommand*{\yXv}{\varSigma}
\newcommand*{\yRv}{\sigma}
\newcommand*{\yxv}{S}
\newcommand*{\yrv}{s}
\newcommand*{\yNv}{\nu}
\newcommand*{\yNN}{\varNu}

\newcommand*{\yx}{\bm{\yxv}}%subpop state
\newcommand*{\yxs}{\sav{\yx}}%subpop av state
\newcommand*{\yX}{\bm{\yXv}}%pop state
\newcommand*{\yXf}{\av{\yX}}%pop av state
\newcommand*{\yxxs}{\sav{\yr\yr}}%subpop av state
\newcommand*{\yXXf}{\av{\yR\yR}}%subpop av state
\newcommand*{\yr}{\bm{\yrv}}%subpop value
\newcommand*{\yrs}{\sav{\yr}}%subpop av value
\newcommand*{\yR}{\bm{\yRv}}%pop value
\newcommand*{\yRf}{\av{\yR}}%pop av value
%conditional assumptions
\newcommand*{\yH}{\varIota}
\newcommand*{\yD}{\varDelta}
\newcommand*{\yHa}{\varIota_\textrm{p}}
\newcommand*{\yHb}{\varIota_\textrm{s}}
\newcommand*{\yHc}{\varKappa}
\newcommand*{\yHd}{\varIota_\textrm{u}}
\newcommand*{\yHp}{\varIota'}
\newcommand*{\yHi}{\varIota''}
%Lagr multipliers
\newcommand*{\yf}{\bm{g}}
\newcommand*{\yc}{\bm{c}}
\newcommand*{\yL}{\bm{\lambda}}
\newcommand*{\yl}{\bm{l}}
\newcommand*{\yk}{z}
\newcommand*{\yK}{\zeta}
%entropies
\newcommand*{\ysh}{H}
\newcommand*{\ybu}{H_{\text{B}}}

@@

Suppose that we have recorded the firing activity of a hundred neurons,
sampled from a particular brain area. What are we to do with such data?
Gerstein, Perkel, Dayhoff \cite{gersteinetal1985} posed this question very
tersely (our emphasis):
\begin{quote}
  The principal conceptual problems are (1) \emph{defining cooperativity or
    functional grouping} among neurons and (2) \emph{formulating
    quantitative criteria} for recognizing and characterizing such
  cooperativity.
\end{quote}
These questions have a long history, of course; see for instance the 1966
review by Moore \etal\
\cite{mooreetal1966}. % This scenario is a concrete possibility
% thanks to recent electrophysiological techniques \cite{berenyietal2014}.
The neuroscientific literature has offered several mathematical definitions
of \enquote{cooperativity} or \enquote{functional grouping} and criteria to
quantify it.

@@

We have already introduced the notation for the sample neurons. We
introduce an analogous notation for the $\yNv$ neurons constituting the
larger population, but using the corresponding Greek letters:
$\yRv_{\iota}(t)$ is the activity of the $\iota$th neuron at time bin $t$,
$\yRf(t) \defd \sum_{\iota} \yRv_{\iota}(t)/\yNv$ is the activity at that
bin averaged over the larger population, and so on.

The probability relations between sample and larger population are valid at
every time bin. As we mentioned above, the \me\ distribution refers to any
time bin or to a new bin. For these reasons we will now omit the time-bin
argument \enquote{(t)} from our expressions.

Probabilities refer to statements about the quantities we observe. We use
the standard notation:
\begin{equation}
  \label{eq:notation_statements}
  \begin{aligned}
    &\text{\enquote{$\yXv_{\iota} = \yRv_{\iota}$} 
      stands for \enquote{the activity of the $\iota$th neuron is $\yRv_{\iota}$}},
    \\
    &\text{\enquote{$\yXf = \yRf$} 
      stands for \enquote{the (population-averaged) activity of the neurons is $\yRf$}},
    \\
    &\text{\enquote{$\yxv_i = \yrv_i$} 
      stands for \enquote{the activity of the $i$th sample neuron is $\yrv_i$}},
  \end{aligned}
\end{equation}
and similarly for other quantities.

If $\yHc$ denotes our state of knowledge -- the evidence and assumptions
backing our probability assignments -- our uncertainty about the full
activity of the larger population is expressed by the joint probability
distribution
\begin{equation}
  \label{eq:joint_plaus}
  \p(\yXv_1=\yRv_1, \yXv_2=\yRv_2, \dotsc, \yXv_\yNv=\yRv_\yNv \| \yHc)
  \quad\text{or}\quad
\p(\yX =\yR \| \yHc), \quad \yR \in \set{0,1}^\yNv.
\end{equation}
Our uncertainty about the state of the sample is likewise expressed by
\begin{equation}
  \label{eq:sample_plaus}
  \p(\yxv_1=\yrv_1, \yxv_2=\yrv_2, \dotsc, \yxv_n=\yrv_n \| \yHc) \quad\text{or}\quad
\p(\yx =\yr \| \yHc), \quad \yr \in \set{0,1}^n.
\end{equation}

@@

The \emph{representation theorem for finite exchangeability} states that,
in the state of knowledge $\yH$, the symmetric distribution for the full
activity is completely determined by the distribution for its
population-average:
\begin{equation}
  \label{eq:joint_plaus_N_homog}
  \p(\yX = \yR \|  \yH) \equiv
  \sum_{\yRf}\p(\yX = \yR \| \yXf = \yRf, \yH)\,
  \p(\yXf=\yRf \| \yH) =
  \binom{\yNv}{\yNv\yRf}^{-1} \p(\yXf=\yRf \| \yH).
\end{equation}
The equivalence on the left is just an application of the law of total
probability; the equality on the right is the statement of the theorem.
This result is intuitive: owing to symmetry, we must assign equal
probabilities to all $\tbinom{\yNv}{\yNv\yRf}$ activity vectors with
$\yNv\yRf$ active neurons; the probability of each activity vector is
therefore given by that of the average activity divided by the number of
possible vector values. Proof of this theorem and generalizations to
non-binary and continuum cases are given by de~Finetti
\cite{definetti1959b}, Kendall \cite{kendall1967}, Ericson
\cite{ericson1976}, Diaconis \amp\ Freedman
\cite{diaconis1977,diaconisetal1980}, Heath \amp\ Sudderth
\cite{heathetal1976}.

Our uncertainties about the full population and the sample are connected
via the conditional probability
\begin{equation}
  \label{eq:conditional_hypergeometric}
  \p(\yxs = \yrs \|\yXf=\yRf, \yH)=
  \binom{n}{n\yrs}\binom{\yNv-n}{\yNv \yRf-n\yrs}\binom{\yNv}{\yNv \yRf}^{-1}
  \defs \ypp(\yrs\|\yRf),
\end{equation}
which is a hypergeometric distribution, typical of \enquote{drawing without
  replacement} problems. The combinatorial proof of this expression is in
fact the same as for this class of problems
\cites[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}.

Using the conditional probability above we obtain the probability for the
activity of the sample:
\begin{gather}
  \label{eq:subpop_average}
  \p(\yxs=\yrs \| \yH) = \sum_{\yRf}
  \p(\yxs = \yrs \|\yXf=\yRf, \yH)\,
  \p(\yXf=\yRf \| \yH)
  = 
  \sum_{\yRf}
  \ypp(\yrs\|\yRf)\,
  \p(\yXf=\yRf \| \yH).
\end{gather}
It should be proved that the probability distribution for the full activity
of the sample is also symmetric and completely determined by the
distribution of its population-averaged activity:
\begin{equation}
  \label{eq:marginal}
  \p(\yx = \yr \| \yH) = \binom{n}{n\yrs}^{-1} \p(\yxs=\yrs \| \yH).
\end{equation}
This is intuitively clear: our initial symmetric ignorance should also
apply to the sample. The distribution for the
sample~\eqref{eq:subpop_average} indeed satisfies the same representation
theorem~\eqref{eq:joint_plaus_N_homog} as the distribution for the full
population.

The conditional probability
$\p(\yxs = \yrs \|\yXf=\yRf, \yH) \equiv \ypp(\yrs \|\yRf)$, besides
relating the distributions for the population and sample activities via
marginalization, also allows us to express the expectation value of any
function of the sample activity, $g(\yrs)$, in terms of the distribution
for the full population, as follows:
\begin{multline}
  \label{eq:pullback_P}
  \expe{g\|I}
  \equiv
  \sum_{\yrs} g(\yrs)\,\p(\yxs=\yrs \| \yH)
  =
  \sum_{\yrs} g(\yrs) \sum_{\yRf} \ypp(\yrs\|\yRf)\,\p(\yXf=\yRf \| \yH)
  ={}\\
  \sum_{\yRf} \biggl[ \sum_{\yrs} g(\yrs)  \ypp(\yrs\|\yRf) \biggr]\,
  \p(\yXf=\yRf \| \yH),
%  =
%  \sum_{N\yRf=0}^N \sum_{n\yrs=0}^n f(\yrs)\ypp(\yrs\|\yRf)\,\p(\yXf=\yRf \| \yH)
%  \equiv  \expeb{g^*}.
\end{multline}
where the second step uses \eqn~\eqref{eq:subpop_average}. The last
expression shows that the expectation of the function $g(\yrs)$ is equal to
the expectation of the function
$g^*(\yRf) \defd \sum_{\yrs} g(\yrs)\,\ypp(\yrs\|\yRf)$.

@@

The final expression in \eqn~\eqref{eq:pullback_P} is important for our
\me\ application: the requirement that the function $g$, defined for the
sample, have a  value $c$ obtained from observed data,
\emph{translates into a linear constraint for the distribution of the full
  population}:
\begin{equation}
  \label{eq:constraint_extended}
  c = \expe{g \| \yH} \equiv \sum_{\yRf} \biggl[ \sum_{\yrs} g(\yrs)  \ypp(\yrs\|\yRf) \biggr]\,
  \p(\yXf=\yRf \| \yH).
\end{equation}

In particular, when the function $g$ is the $m$-activity of the sample,
$g(\yrs) = \sav{\yr\dotso\yr} \equiv \binom{n\yrs}{m}/\binom{n}{m}$, we
find
\begin{equation}
  \label{eq:expe_products}
%  \label{eq:pullback_m_expectations}
%  \ypp^*\colon \sav{\underbrace{\yx \dotsm \yx}_{\text{$m$ factors}}}
%  \mapsto
    % \bigl( \sav{\underbrace{\yx \dotsm \yx}_{\text{$m$ factors}}}\bigr)^* =
    % \av{\underbrace{\yX \dotsm \yX}_{\text{$m$ factors}}},
    % \\[2\jot]
    % \expe{\sav{\underbrace{\yx \dotsm \yx}_{\text{$m$ factors}}} \| \yH} =
    % \expe{\av{\underbrace{\yX \dotsm \yX}_{\text{$m$ factors}}} \| \yH}
  \expe{\sav{\underbrace{\yr \dotsm \yr}_{\text{$m$ factors}}} \| \yH}
\equiv
    % \\[2\jot]
    % {}
    % \frac{(n-m)!}{n!}
    \sum_{\yrs} %m!\,
    \binom{n}{m}^{-1}
    \binom{n \yrs}{m}\, \p(\yxs=\yrs \| \yH)
    % \sum_{N \yRf=0}^N %m!\,
    % \binom{N-m}{N \yRf-m} \binom{N}{N \yRf}^{-1} \p(\yXf=\yRf \| \yH)
    = \binom{\yNv}{m}^{-1}
    % \frac{(N-m)!}{N!}
    \sum_{\yRf} %m!\,
    \binom{\yNv \yRf}{m}\, \p(\yXf=\yRf \| \yH)
\equiv    \expe{\av{\underbrace{\yR \dotsm \yR}_{\text{$m$ factors}}} \| \yH},
\end{equation}
that is, \emph{the expected values of the $m$-activities of the sample and
  of the full population are equal}. The proof of the middle equality uses
the expression for the $m$th factorial moment of the hypergeometric
distribution and can be found in \textcite{potts1953}. Similar relations
can be found for the raw moments $\expe{\yrs^m}$ and $\expe{\yRf^m}$, which
can be written in terms of the product expectations using
\eqn~\eqref{eq:products_intermsof_average}.

@@

\begin{figure}[!b]
\centering
\includegraphics[width=0.9\linewidth]{pop_sample_projection3.pdf}%
\caption{Log-density plot of the hypergeometric distribution
  $%\p(\yxs = \yrs \|\yXf=\yRf, \yH)=
\ypp(\yrs\|\yRf) \defd  \raisebox{0pt}[0pt][1ex]{}\binom{n}{n\yrs}\binom{\yNv-n}{\yNv \yRf-n\yrs}\binom{\yNv}{\yNv \yRf}^{-1}$ for $\yNv=5000$, $n=200$. (Band artifacts may appear in the
  colourbar depending on your \textsc{pdf} viewer.)}
\label{fig:hypergeom_proj}
\end{figure}%hypergeometric_identity.nb
These expectation equalities between sample and full population should not
be surprising: we intuitively \emph{expect} that the proportion of coloured
balls sampled from an urn should be roughly equal to the proportion of
coloured ball contained in the urn. The formulae in the present section
formalize and mathematically express our intuition. The hypergeometric
distribution $\ypp(\yrs\|\yRf)$ plays an important role in this
formalization. A look at its plot, \fig~\ref{fig:hypergeom_proj}, reveals
that it is a sort of \enquote{fuzzy identity transformation}, or fuzzy
Kronecker delta, between the $\yRf$-space $\set{0,\dotsc,\yNv}$ and
$\yrs$-space $\set{0,\dotsc,n}$. From \eqn~\eqref{eq:marginal} we thus have
that
\begin{equation}
  \label{eq:roughly_equal_nN}
  \p(\yxs=a \|\yH) \approx \p(\yXf=a \|\yH),\qquad
\expeb{g(\yrs) \|\yH} \approx \expeb{g(\yRf) \|\yH},
\end{equation}
where $g$ is any smooth function defined on $\clcl{0,1}$. These approximate
equalities express the intuitive fact that \emph{our uncertainty about the
  sample is representative of our uncertainty about the population and
  about other samples}, and vice versa. When $n=\yNv$, $\ypp(\yrs\|\yRf)$
becomes the identity matrix and the approximate equalities above become
exact -- of course, since we have sampled the full population.

@@

Applied directly to the sampled neurons, the method yields the distribution
\begin{equation}
  \label{eq:app_maxent_sample}
  \p(\yxs=\yrs \|\yHb)
  =\frac{1}{\yk(\yl)}\,
  \binom{n}{n\yrs}
  \exp[\yl\T \yf(\yrs)]
\end{equation}
where $\yk(\yl)$ is a normalization constant. The binomial in front of the
exponential appears because we must account for the multiplicity by which
the population-average activity $\yrs$ can be realized: $\yrs=0$ can be
realized in only one way (all neurons inactive), $\yrs=1/n$ can be realized
in $n$ ways (one active neuron out of $n$), and so on. This term is analogous to
the \enquote{density of states} in front of the Boltzmann factor in
statistical mechanics \cite[\chap~16]{callen1960_r1985}. The $m$ Lagrange
multipliers $\yl\defd (l_1,\dotsc,l_m)$ must satisfy the $m$ constraint
equations
\begin{equation}
  \label{eq:app_maxent_sample_constraints}
  c_k = \expe{g_k \|\yHb} \equiv
  \frac{1}{\yk(\yl)}\sum_{\yrs}  g_k(\yrs) \binom{n}{n\yrs}
  \exp[\yl\T \yf(\yrs)], \qquad k=1,\dotsc,m.
\end{equation}

Applied to the full population, using the constraint
expression~\eqref{eq:constraint_extended} derived in the previous section,
the method yields the distribution for the full-population activity
\begin{equation}
  \label{eq:app_maxent_pop}
  \p(\yXf = \yRf \| \yHa)  = \frac{1}{\yK(\yL)}\,
  \binom{\yNv}{\yNv \yRf}\,\exp\Bigl[\yL\T
  \sum_{\yrs} \yf(\yrs)\ypp(\yrs\|\yRf)\Bigr].
\end{equation}
The $m$ Lagrange multipliers $\yL\defd (\lambda_1,\dotsc,\lambda_m)$ must
satisfy the $m$ constraint equations
\begin{equation}
  \label{eq:app_maxent_pop_constraints}
  c_k = \expe{g_k \|\yHa}\equiv
  \frac{1}{\yK(\yL)}\sum_{\yrs} \sum_{\yRf} g_k(\yrs) \ypp(\yrs \|\yRf)
\,
  \binom{\yNv}{\yNv \yRf}\,\exp\Bigl[\yL\T
  \sum_{\yrs} \yf(\yrs)\ypp(\yrs\|\yRf)\Bigr],
  \qquad k=1,\dotsc,m.
\end{equation}

We obtain the distribution for the sample activity by marginalization, using
\eqn~\eqref{eq:marginal}:
\begin{equation}
  \label{eq:app_maxent_pop_marg}
  \p(\yxs = \yrs \| \yHa)  = \frac{1}{\yK(\yL)}\, 
  \sum_{\yRf} \ypp(\yrs \|\yRf)
\,
  \binom{\yNv}{\yNv \yRf}\,\exp\Bigl[\yL\T
  \sum_{\yrs} \yf(\yrs)\ypp(\yrs\|\yRf)\Bigr].
\end{equation}

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-11-04T09:20:08+0100
%% Last-Updated: 2018-12-18T13:21:16+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Report-no: ***
\documentclass{article}

@@

% \DeclareMathSymbol{\iota}{\mathalpha}{egreeki}{"13}

@@

\newcommand*{\p}{\mathrm{p}}%probability

@@

%\newcommand*{\yXv}{\varSigma}
\newcommand*{\yRv}{S}
%\newcommand*{\yxv}{s}
\newcommand*{\yrv}{s}
\newcommand*{\yNv}{N}
\newcommand*{\yNN}{\varNu}

%\newcommand*{\yx}{\bm{\yxv}}%subpop state
%\newcommand*{\yxs}{\sav{\yx}}%subpop av state
%\newcommand*{\yX}{\bm{\yXv}}%pop state
%\newcommand*{\yXf}{\av{\yX}}%pop av state
\newcommand*{\yrrs}{\sav{\yr\yr}}%subpop av state
%\newcommand*{\yXXf}{\av{\yR\yR}}%subpop av state
\newcommand*{\yr}{\bm{\yrv}}%subpop value
\newcommand*{\yrs}{\sav{\yr}}%subpop av value
\newcommand*{\yR}{\bm{\yRv}}%pop value
\newcommand*{\yRf}{\av{\yR}}%pop av value
%conditional assumptions
\newcommand*{\yH}{\varIota}
\newcommand*{\yD}{\varDelta}
\newcommand*{\yHa}{\varIota_\textrm{p}}
\newcommand*{\yHb}{\varIota_\textrm{s}}
\newcommand*{\yHc}{\varKappa}
\newcommand*{\yHd}{\varIota_\textrm{u}}
\newcommand*{\yHp}{\varIota'}
\newcommand*{\yHi}{\varIota''}
%Lagr multipliers
\newcommand*{\yf}{\bm{g}}
\newcommand*{\yc}{\bm{c}}
\newcommand*{\yL}{\bm{\lambda}}
\newcommand*{\yl}{\bm{l}}
\newcommand*{\yk}{z}
\newcommand*{\yK}{\zeta}
%entropies
\newcommand*{\ysh}{H}
\newcommand*{\ybu}{H_{\text{B}}}

@@

Suppose we have recorded the firing activity of a hundred neurons,
sampled from a particular brain area. What are we to do with such data?
Gerstein, Perkel, Dayhoff \cite{gersteinetal1985} posed this question very
tersely (our emphasis):
\begin{quote}
  The principal conceptual problems are (1) \emph{defining cooperativity or
    functional grouping} among neurons and (2) \emph{formulating
    quantitative criteria} for recognizing and characterizing such
  cooperativity.
\end{quote}
These questions have a long history, of course; see for instance the 1966
review by Moore \etal\
\cite{mooreetal1966}. % This scenario is a concrete possibility
% thanks to recent electrophysiological techniques \cite{berenyietal2014}.
The neuroscientific literature has offered several mathematical definitions
of \enquote{cooperativity} or \enquote{functional grouping} and criteria to
quantify it.

@@

We have already introduced the notation for the sample neurons. We
introduce an analogous notation for the $\yNv$ neurons constituting the
larger population, but using the corresponding capital letters:
$\yRv_{i}(t)$ is the activity of the $i$th neuron at time bin $t$,
$\yRf(t) \defd \sum_{i} \yRv_{i}(t)/\yNv$ is the activity at that
bin averaged over the larger population, and so on.

The probability relations between sample and larger population are valid at
every time bin. As we mentioned above, the \me\ distribution refers to any
time bin or to a new bin. For these reasons we will now omit the time-bin
argument \enquote{$(t)$} from our expressions.

% Probabilities refer to statements about the quantities we observe. We use
% the standard notation:
% \begin{equation}
%   \label{eq:notation_statements}
%   \begin{aligned}
%     &\text{\enquote{$\yXv_{\iota} = \yRv_{\iota}$} 
%       stands for \enquote{the activity of the $\iota$th neuron is $\yRv_{\iota}$}},
%     \\
%     &\text{\enquote{$\yXf = \yRf$} 
%       stands for \enquote{the (population-averaged) activity of the neurons is $\yRf$}},
%     \\
%     &\text{\enquote{$\yxv_i = \yrv_i$} 
%       stands for \enquote{the activity of the $i$th sample neuron is $\yrv_i$}},
%   \end{aligned}
% \end{equation}
% and similarly for other quantities.

If $\yHc$ denotes our state of knowledge -- the evidence and assumptions
backing our probability assignments -- our uncertainty about the full
activity of the larger population is expressed by the joint probability
distribution
\begin{equation}
  \label{eq:joint_plaus}
  \p(\yRv_1, \yRv_2, \dotsc, \yRv_\yNv \| \yHc)
  \quad\text{or}\quad
\p(\yR \| \yHc), \quad \yR \in \set{0,1}^\yNv.
\end{equation}
Our uncertainty about the state of the sample is likewise expressed by
\begin{equation}
  \label{eq:sample_plaus}
  \p(\yrv_1, \yrv_2, \dotsc, \yrv_n \| \yHc) \quad\text{or}\quad
\p(\yr \| \yHc), \quad \yr \in \set{0,1}^n.
\end{equation}

@@

The \emph{representation theorem for finite exchangeability} states that,
in the state of knowledge $\yH$, the symmetric distribution for the full
activity is completely determined by the distribution for its
population-average:
\begin{equation}
  \label{eq:joint_plaus_N_homog}
  \p(\yR \|  \yH) \equiv
  \sum_{\yRf}\p(\yR \| \yRf, \yH)\,
  \p(\yRf \| \yH) =
  \binom{\yNv}{\yNv\yRf}^{-1} \p(\yRf \| \yH).
\end{equation}
The equivalence on the left is just an application of the law of total
probability; the equality on the right is the statement of the theorem.
This result is intuitive: owing to symmetry, we must assign equal
probabilities to all $\tbinom{\yNv}{\yNv\yRf}$ activity vectors with
$\yNv\yRf$ active neurons; the probability of each activity vector is
therefore given by that of the average activity divided by the number of
possible vector values. Proof of this theorem and generalizations to
non-binary and continuum cases are given by de~Finetti
\cite{definetti1959b}, Kendall \cite{kendall1967}, Ericson
\cite{ericson1976}, Diaconis \amp\ Freedman
\cite{diaconis1977,diaconisetal1980}, Heath \amp\ Sudderth
\cite{heathetal1976}.

Our uncertainties about the full population and the sample are connected
via the conditional probability
\begin{equation}
  \label{eq:conditional_hypergeometric}
  \p(\yrs \|\yRf, \yH)=
  \binom{n}{n\yrs}\binom{\yNv-n}{\yNv \yRf-n\yrs}\binom{\yNv}{\yNv \yRf}^{-1}
  \defs \ypp(\yrs\|\yRf),
\end{equation}
which is a hypergeometric distribution, typical of \enquote{drawing without
  replacement} problems. The combinatorial proof of this expression is in
fact the same as for this class of problems
\cites[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}.

Using the conditional probability above we obtain the probability for the
activity of the sample:
\begin{gather}
  \label{eq:subpop_average}
  \p(\yrs \| \yH) = \sum_{\yRf}
  \p(\yrs \|\yRf, \yH)\,
  \p(\yRf \| \yH)
  = 
  \sum_{\yRf}
  \ypp(\yrs\|\yRf)\,
  \p(\yRf \| \yH).
\end{gather}
It should be proved that the probability distribution for the full activity
of the sample is also symmetric and completely determined by the
distribution of its population-averaged activity:
\begin{equation}
  \label{eq:marginal}
  \p(\yr \| \yH) = \binom{n}{n\yrs}^{-1} \p(\yrs \| \yH).
\end{equation}
This is intuitively clear: our initial symmetric ignorance should also
apply to the sample. The distribution for the
sample~\eqref{eq:subpop_average} indeed satisfies the same representation
theorem~\eqref{eq:joint_plaus_N_homog} as the distribution for the full
population.

The conditional probability
$\p(\yrs \|\yRf, \yH) \equiv \ypp(\yrs \|\yRf)$, besides
relating the distributions for the population and sample activities via
marginalization, also allows us to express the expectation value of any
function of the sample activity, $g(\yrs)$, in terms of the distribution
for the full population, as follows:
\begin{equation}
  \label{eq:pullback_P}
  \expe{g\|I}
  \equiv
  \sum_{\yrs} g(\yrs)\,\p(\yrs \| \yH)
  =
  \sum_{\yrs} g(\yrs) \sum_{\yRf} \ypp(\yrs\|\yRf)\,\p(\yRf \| \yH)
  = %{}\\
  \sum_{\yRf} \biggl[ \sum_{\yrs} g(\yrs)  \ypp(\yrs\|\yRf) \biggr]\,
  \p(\yRf \| \yH),
%  =
%  \sum_{N\yRf=0}^N \sum_{n\yrs=0}^n f(\yrs)\ypp(\yrs\|\yRf)\,\p(\yXf=\yRf \| \yH)
%  \equiv  \expeb{g^*}.
\end{equation}
where the second step uses \eqn~\eqref{eq:subpop_average}. The last
expression shows that the expectation of the function $g(\yrs)$ is equal to
the expectation of the function
$g^*(\yRf) \defd \sum_{\yrs} g(\yrs)\,\ypp(\yrs\|\yRf)$.

@@

The final expression in \eqn~\eqref{eq:pullback_P} is important for our
\me\ application: the requirement that the function $g$, defined for the
sample, have a  value $c$ obtained from observed data,
\emph{translates into a linear constraint for the distribution of the full
  population}:
\begin{equation}
  \label{eq:constraint_extended}
  c = \expe{g \| \yH} \equiv \sum_{\yRf} \biggl[ \sum_{\yrs} g(\yrs)  \ypp(\yrs\|\yRf) \biggr]\,
  \p(\yRf \| \yH).
\end{equation}

In particular, when the function $g$ is the $m$-activity of the sample,
$g(\yrs) = \sav{\yr\dotso\yr} \equiv \binom{n\yrs}{m}/\binom{n}{m}$, we
find
\begin{equation}
  \label{eq:expe_products}
%  \label{eq:pullback_m_expectations}
%  \ypp^*\colon \sav{\underbrace{\yx \dotsm \yx}_{\text{$m$ factors}}}
%  \mapsto
    % \bigl( \sav{\underbrace{\yx \dotsm \yx}_{\text{$m$ factors}}}\bigr)^* =
    % \av{\underbrace{\yX \dotsm \yX}_{\text{$m$ factors}}},
    % \\[2\jot]
    % \expe{\sav{\underbrace{\yx \dotsm \yx}_{\text{$m$ factors}}} \| \yH} =
    % \expe{\av{\underbrace{\yX \dotsm \yX}_{\text{$m$ factors}}} \| \yH}
  \expe{\sav{\underbrace{\yr \dotsm \yr}_{\text{$m$ factors}}} \| \yH}
\equiv
    % \\[2\jot]
    % {}
    % \frac{(n-m)!}{n!}
    \sum_{\yrs} %m!\,
    \binom{n}{m}^{-1}
    \binom{n \yrs}{m}\, \p(\yrs \| \yH)
    % \sum_{N \yRf=0}^N %m!\,
    % \binom{N-m}{N \yRf-m} \binom{N}{N \yRf}^{-1} \p(\yXf=\yRf \| \yH)
    = \binom{\yNv}{m}^{-1}
    % \frac{(N-m)!}{N!}
    \sum_{\yRf} %m!\,
    \binom{\yNv \yRf}{m}\, \p(\yRf \| \yH)
\equiv    \expe{\av{\underbrace{\yR \dotsm \yR}_{\text{$m$ factors}}} \| \yH},
\end{equation}
that is, \emph{the expected values of the $m$-activities of the sample and
  of the full population are equal}. The proof of the middle equality uses
the expression for the $m$th factorial moment of the hypergeometric
distribution and can be found in \textcite{potts1953}. Similar relations
can be found for the raw moments $\expe{\yrs^m}$ and $\expe{\yRf^m}$, which
can be written in terms of the product expectations using
\eqn~\eqref{eq:products_intermsof_average}.

@@

\begin{figure}[!b]
\centering
\includegraphics[width=0.9\linewidth]{pop_sample_projection3.pdf}%
\caption{Log-density plot of the hypergeometric distribution
  $%\p(\yrs \|\yRf, \yH)=
\ypp(\yrs\|\yRf) \defd  \raisebox{0pt}[0pt][1ex]{}\binom{n}{n\yrs}\binom{\yNv-n}{\yNv \yRf-n\yrs}\binom{\yNv}{\yNv \yRf}^{-1}$ for $\yNv=5000$, $n=200$. (Band artifacts may appear in the
  colourbar depending on your \textsc{pdf} viewer.)}
\label{fig:hypergeom_proj}
\end{figure}%hypergeometric_identity.nb
These expectation equalities between sample and full population should not
be surprising: we intuitively \emph{expect} that the proportion of coloured
balls sampled from an urn should be roughly equal to the proportion of
coloured ball contained in the urn. The formulae in the present section
formalize and mathematically express our intuition. The hypergeometric
distribution $\ypp(\yrs\|\yRf)$ plays an important role in this
formalization. A look at its plot, \fig~\ref{fig:hypergeom_proj}, reveals
that it is a sort of \enquote{fuzzy identity transformation}, or fuzzy
Kronecker delta, between the $\yRf$-space $\set{0,\dotsc,\yNv}$ and
$\yrs$-space $\set{0,\dotsc,n}$. From \eqn~\eqref{eq:marginal} we thus have
that
\begin{equation}
  \label{eq:roughly_equal_nN}
  \p(\yrs=a \|\yH) \approx \p(\yRf=a \|\yH),\qquad
\expeb{g(\yrs) \|\yH} \approx \expeb{g(\yRf) \|\yH},
\end{equation}
where $g$ is any smooth function defined on $\clcl{0,1}$. These approximate
equalities express the intuitive fact that \emph{our uncertainty about the
  sample is representative of our uncertainty about the population and
  about other samples}, and vice versa. When $n=\yNv$, $\ypp(\yrs\|\yRf)$
becomes the identity matrix and the approximate equalities above become
exact -- of course, since we have sampled the full population.

@@

Applied directly to the sampled neurons, the method yields the distribution
\begin{equation}
  \label{eq:app_maxent_sample}
  \p(\yrs \|\yHb)
  =\frac{1}{\yk(\yl)}\,
  \binom{n}{n\yrs}
  \exp[\yl\T \yf(\yrs)]
\end{equation}
where $\yk(\yl)$ is a normalization constant. The binomial in front of the
exponential appears because we must account for the multiplicity by which
the population-average activity $\yrs$ can be realized: $\yrs=0$ can be
realized in only one way (all neurons inactive), $\yrs=1/n$ can be realized
in $n$ ways (one active neuron out of $n$), and so on. This term is analogous to
the \enquote{density of states} in front of the Boltzmann factor in
statistical mechanics \cite[\chap~16]{callen1960_r1985}. The $m$ Lagrange
multipliers $\yl\defd (l_1,\dotsc,l_m)$ must satisfy the $m$ constraint
equations
\begin{equation}
  \label{eq:app_maxent_sample_constraints}
  c_k = \expe{g_k \|\yHb} \equiv
  \frac{1}{\yk(\yl)}\sum_{\yrs}  g_k(\yrs) \binom{n}{n\yrs}
  \exp[\yl\T \yf(\yrs)], \qquad k=1,\dotsc,m.
\end{equation}

Applied to the full population, using the constraint
expression~\eqref{eq:constraint_extended} derived in the previous section,
the method yields the distribution for the full-population activity
\begin{equation}
  \label{eq:app_maxent_pop}
  \p(\yRf \| \yHa)  = \frac{1}{\yK(\yL)}\,
  \binom{\yNv}{\yNv \yRf}\,\exp\Bigl[\yL\T
  \sum_{\yrs} \yf(\yrs)\ypp(\yrs\|\yRf)\Bigr].
\end{equation}
The $m$ Lagrange multipliers $\yL\defd (\lambda_1,\dotsc,\lambda_m)$ must
satisfy the $m$ constraint equations
\begin{equation}
  \label{eq:app_maxent_pop_constraints}
  c_k = \expe{g_k \|\yHa}\equiv
  \frac{1}{\yK(\yL)}\sum_{\yrs} \sum_{\yRf} g_k(\yrs) \ypp(\yrs \|\yRf)
\,
  \binom{\yNv}{\yNv \yRf}\,\exp\Bigl[\yL\T
  \sum_{\yrs} \yf(\yrs)\ypp(\yrs\|\yRf)\Bigr],
  \qquad k=1,\dotsc,m.
\end{equation}

We obtain the distribution for the sample activity by marginalization, using
\eqn~\eqref{eq:marginal}:
\begin{equation}
  \label{eq:app_maxent_pop_marg}
  \p(\yrs \| \yHa)  = \frac{1}{\yK(\yL)}\, 
  \sum_{\yRf} \ypp(\yrs \|\yRf)
\,
  \binom{\yNv}{\yNv \yRf}\,\exp\Bigl[\yL\T
  \sum_{\yrs} \yf(\yrs)\ypp(\yrs\|\yRf)\Bigr].
\end{equation}

The distributions for the sample activity,
\eqns~\eqref{eq:app_maxent_pop_marg} and \eqref{eq:app_maxent_sample},
obtained with the two approaches $\yHb$ and $\yHa$, are different. From the
discussion in the previous section we expect them to be vaguely similar;
yet they cannot be exactly equal, because their equality would require the
$2m$ quantities $\yL$ and $\yl$ to satisfy the constraint equations
\eqref{eq:app_maxent_pop_constraints} and
\eqref{eq:app_maxent_sample_constraints}, and in addition also the $n$
equations $\p(\yrs \| \yHa) = \p(\yrs \| \yHb)$,
$\yrs=1/n,\dotsc,1$ (one equation is taken care of by the normalization of
the distributions). We would have a set of $2m+n$ equations in $2m$
unknowns.
